%--------------------------------------------------------------------------------------------------
% 
\chapter*{Acknowledgments}
\pdfbookmark[0]{Acknowledgments}{Acknowledgments}
%--------------------------------------------------------------------------------------------------

I would like to take this opportunity to express my deep gratitude and appreciation to the many people who have supported and guided me throughout my doctoral journey.

First and foremost, I am profoundly thankful to my advisor, prof. dr. Dunja Mladenić, for her unwavering support, invaluable guidance, and patience throughout the entire research process. 
Her mentorship and expertise have been instrumental in shaping this work and my academic growth.

I also thank the members of my dissertation committee, prof. dr. Chrysi Laspidou, prof. dr. Aleš Švigelj, and dr. Blaž Fortuna, for their insightful feedback and critical evaluation of my research.
Their collective wisdom has greatly enriched the quality of this thesis.

I am grateful to my colleagues and fellow researchers at Jožef Stefan Institute, Erik Novak, dr. Alenka Guček, Maja Škrjanc, Filip Koprivec, Zala Herga Rupnik, prof. dr. Primož Škraba, Marko Grobelnik, dr. Inna Novalja, dr. Jože Rožanec and Jože Peternelj for their stimulating discussions, collaboration, camaraderie and insightful debates that so often exceeded the informal limitations of the work environment.
New horizons have been revealed to me also by the knowledgeable coworkers at the Qlector d.o.o., among them I especially thank Viktor Jovanoski, dr. Jan Rupnik, and dr. Blaž Kažič.
Their contributions have played a significant role in shaping my ideas and refining my research.
I want to express my sincere thanks to our office staff, Jasna Franko and Mojca Kregar.
Their practical assistance, patience, and emotional support have been invaluable throughout this project.
Their dedication and willingness to go the extra mile have been greatly appreciated.

% projects
I would also like to thank numerous coworkers on several research projects through my years at the Jožef Stefan Institute. 
Especially all the researchers and students who shaped the work and ideas in FP7 NRG4CAST, FP7 Sunseed, H2020 Water4Cities, H2020 NAIDES, and H2020 PerceptiveSentinel projects.
All of them can not be mentioned here but have in various ways contributed to the successful completion of this thesis. 

Lastly, I want to keep a special place in these acknowledgments for my friend Grega Milčinski.
He helped to kick off my career after enrolling in the PhD program by inviting me to collaborate in H2020 PerceptiveSentinel proposal, which became my first of several accepted proposals in the European Comission research calls.
Moreover, he always offered me a view from afar of the problems I faced during this time and helped me with constructive solutions.

%Lastly, I reserve a special section in these acknowledgments for my friend, Grega Milčinski. 
%Grega played a crucial role in launching my academic career when, upon my enrollment in the PhD program, he invited me to collaborate on the H2020 PerceptiveSentinel proposal. 
%This collaboration marked the beginning of a successful journey, leading to the acceptance of several research proposals in Horizon 2020 and Horizon Europe calls.
%Beyond that, when needed, Grega offered me an invaluable outside perspective on the difficulties I experienced during this time, offering helpful insights and encouragement.